---
title: "Reproducible Research: Peer Assessment 1"
by: Syaiful
output: 
  html_document:
    keep_md: true
---
This assignment makes use of data from a personal activity monitoring device. This device collects data at 5 minute intervals through out the day. The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day.

The data for this assignment can be downloaded from the course web site:

* Dataset: [Activity monitoring data](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip)

## Loading and preprocessing the data
Load the package
Read the data from your working directory to obtain a csv file.

```{r}
library("data.table")
library(ggplot2)

activitydata <- data.table::fread(input = "activity.csv")
```

## What is mean total number of steps taken per day?
* number of steps taken per day

```{r}
Total_Steps <- activitydata[, c(lapply(.SD, sum, na.rm = FALSE)), .SDcols = c("steps"), by = .(date)] 
head(Total_Steps, 10)
```
* Histogram of the total number of steps taken each day 

```{r}
ggplot(Total_Steps, aes(x = steps)) +
    geom_histogram(binwidth = 1000) +
    labs(title = "Daily Steps", x = "Steps", y = "Frequency")+
  theme_classic()
```


* mean and median of steps taken per day
```{r}
Total_Steps[, .(Mean_Steps = mean(steps, na.rm = TRUE), Median_Steps = median(steps, na.rm = TRUE))]
```

## What is the average daily activity pattern?
* Time series plot of the average number of steps taken

```{r}
IntervalDT <- activitydata[, c(lapply(.SD, mean, na.rm = TRUE)), .SDcols = c("steps"), by = .(interval)] 
ggplot(IntervalDT, aes(x = interval , y = steps)) + geom_line(color="black", size=1) + labs(title = "Average Daily Steps", x = "Interval", y = "Avg. Steps per day") + theme_classic()
```


* The 5-minute interval that, on average, contains the maximum number of steps

```{r}
IntervalDT[steps == max(steps), .(max_interval = interval)]
```


## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
